---
title: "How film length and budget are positively related to a film's reception, while year of release is negatively related to a film's reception"
author: "Arjun Dhatt"
date: "20/12/2020"
output:
  pdf_document: default
  html_document: default
---

## Keywords 
- Film
- Year
- Budget 
- Linear Regression 
- Title 

## Abstract

In this report, I will be looking at the various factors that influence a film's reception among audiences. Using the 'ggplot2movies' packages, I will be looking at the dataset 'movies' and filtering the dataset to include relevant variables such as — title, rating, length, year, and budget. I will then run an analysis to identify whether the variables 'budget', 'length', or 'year' affect a film's reception. After completing the analysis, I concluded that a film's length and budget have a positive relationship with the film's rating, while year has a negative relationship. 

## Introduction 

Many variables can influence an audience's reception to a film. These variables include— the budget of the film, the length of the film, and the year the film was released. This poses the interesting question — how does a film's budget, length and year of release affect it's reception among audiences?

In the report, I will investigate this question by taking a deep look at the interplay between all the factors. I began the analysis by cleaning the 'ggplot2movies' data so that it only includes variables of interest such as — title, length, rating, budget, and year of release. Since some of these entries in our dataset do not contain information for relevant variables, I deleted those entries so that all films in the dataset are analyzed from the same standpoint. 

I began the analysis by looking at the variables budget and length. I conducted linear regression for both variables and discovered that there is a significant positive relationship due to the significantly small p-value. Lastly, I looked at the variable year and discovered  that there is a significant negative relationship due to the significantly small p-value. With these results, I concluded that a film's budget and length positively affect a film's reception, while a film's year of release negatively affect a film's reception.

The results from this analysis are important because they provide an outlook on important variables that affect a film's reception. This information can be useful to both audiences that watch films as well as to film production companies so they can forecast what the reception for their films will be like. 

In this report, I will begin by explaining the data that I will be using in the 'Data' section, followed by the model I will use to make conclusions in the 'Model' section. I will then show the results from my analysis in the 'Results' section and explain them in the 'Discussion' sections. Lastly, I will reserve a section for any extranneous information, code, and references in the 'Appendix/References' section.


## Data

```{r}
library(ggplot2movies)
library(ggplot2)
library(dplyr)
movies <- mutate(movies, decade = ifelse(year %in% 1950:1959, "1950s",
                                        ifelse(year %in% 1960:1969, "1960s",
                                        ifelse(year %in% 1970:1979, "1970s",
                                        ifelse(year %in% 1980:1989, "1980s",
                                        ifelse(year %in% 1990:1999, "1990s",
                                        ifelse(year %in% 2000:2009, "2000s", "<1949")))))))

movies <- mutate(movies, )

moviesx <-movies[c('title', 'year', 'length', 'budget','rating','votes') ]


moviesfiltered <- filter(moviesx, votes > (quantile(movies$votes, .5)))
moviesfilteredx <- filter(moviesfiltered, year > 1950)

```

In this report, I will be using the ‘ggplot2movies’ dataset. The data is collected from the website https://www.imdb.com — a website dedicated to collecting information about various films, television shows, and celebrities. Users are able to complete reviews on the IMDB website allowing them to provide a score for a film on a scale from 1-10. 

The ‘ggplot2movies’ package selects films from the IMDB site if they had a known length and had been rated by at least one IMDB user. Each film corresponds to one entry in the dataset and contains information on up to 25 variables; not all films contain information about all 25 categories. For example, many of the films from before 1960 do not contain information about the film budget which will be important when determining whether budget affects a films reception. The ‘ggplot2movies’ package contains information on a total of 58,788 films. All of the films used in the dataset were collected from the year 1893 to 2005. 

Although the dataset contains information on 25 variables, I will only be analyzing a select few variables in the dataset. The variables and descriptions of the variables I will be analyzing can be seen below:

-	‘title’ — provides the titles of the movie
-	‘year’ — provides the year that the film was released 
-	‘budget’ — provides the total budget of the film in US dollars 
-	‘length’ — provides the length of the film
-	‘rating’ — provides the average rating of the film by IMDB users 

A large part of this report will use rating as a dependant variable and rating is largely dependant on the number of votes. The original dataset contains 58,788 films, but many of those films contain very little votes. According to the law of large number, as the sample size grows larger, the rating will approach its true mean; the rating can not approach the true mean if it is based on only a few number of votes. Therefore, I removed all entries with less than 30 votes. The 50 vote threshold was based on the 50th percentile of 'votes'. Removing al entries below the 50th percentile ensures that the rating is not unreliable due to a few number of votes. 

When observing a histogram based on year of release, we can see that the dataset is severely left-skewed:

```{r}
library(ggplot2movies)
library(dplyr)
hist(movies$year)
```

Many entries from the data are from before the year 1950; the dataset actually contains 10,418 entries from films before 1950. Films from before 1950 are different from the films that we are used to seeing in today's age in many different ways. Reviewers may have focused on stylistic traits, rather than both stylistic and narrative traits like we see in today's movies; movies were also shorter. Due to the unfair ground we may be measuring the films from different eras, I decided to focus exclusively on films released after 1950's as they conform to the style we are used to seeing in modern films. 

To make the variable ‘year’ easier to analyze, I created a variable titled ‘decade’ so that the films are grouped by decade. This makes it much easier to analyze because there are fewer values to analyze. 

Below, you can see what the first few values of our dataset looklike:

```{r}
head(moviesfilteredx)
```

The population in this case is every person that watched the film and can rate the film on a scale from 1 to 10. The population frame in this case is all people who have an IMDB account, watched the film, and can rate the film on a scale from 1 to 10. Entries in the dataset were found by searching through IMDB and collecting data on films if they had a known length and had been rated by at least one IMDB user. 

The key features of the data is the breadth of films that the data offers. We get information about films from various eras of cinema ensuring that our data is a good representation of films in general. The advantage of the dataset arises from the fact that our data is well representative of film in general; this means that when we perform an analysis on the data, we know that our results are meaningful. Despite the fact that it is well representative of film, it's largely representative of Western cinema and missing out other large industries around the world; this can be viewed as a disadvantage because our data does not represent the entire film industry. If our data included films from the Eastern world, the results from our analysis may be different.

GO MORE IN DEPTH INTO THE BIASES OF THE DATA


## Model

In this paper, I will be using linear regression to see how different variables can affect a films reception. 

A linear regression model can be explained by the equation $$y=a*x_0 +b$$. In this case, $y$ is the film's dependant variable, while $x_0$ is the independant variable that I want to determine if it affects the dependant variable. Using the lm() function in R, I will create a linear regression model, and then I will look at the summary of the model to see if the p-value is significant. If the p-value is less than 0.05, then I know that the independant variable affects the films reception. 

The dependant variable in this linear regression model will be 'rating' because I want to see how the rating of a film is dependant on other independant variable. Since I am analyzing a 'film's reception' in this paper, the variable that best encompasses reception is the average rating that IMDB users provide for the film. 

The independant variables I will be analyzing are: 
- 'length'- this is simply provided by the variable 'length'
- 'year'- this is simply provided by the variable 'year'. I did mutate another variable titled 'decade' which makes visualizing the data easier, but when performing linear regression, 'year' is more thorough so I decided to use 'year' as an independant variable. 
- 'budget'- this is simply provided by the variable 'budget'

All 3 variables that I am analyzing are continuous variables. 

Because I am identifying how a film's reception is affected by various independant variables, it makes sense to run linear regression as I want to see the cause and effect relationship between an independant and dependant variable. The relationships that I am analyzing are strictly between 2 variables — rating (dependant) and length, year or budget (independant). 

## Results 

In this section of this paper, I will be looking at the relation between a film's rating and 3 different independant variables— length, year and budget. I will be using a signicance level of 0.05. 

### How budget affects a film's reception 

I will first look at how a film's budget affects the reception of a film

In order to determine the relationship between the dependant variable (reception) and the independant variable (budget), I will run a linear regression. Because many of the entries in our dataset don't contain information about the film's budget, I will remove those values when performing the linear regression; this means that the linear regression between budget and reception will only inspect entries with a budget greater than 0. 

```{r}
#Scatterplot of rating and budget 

#Remove entries that have a value of NA or 0 for budget 
budgetfiltered <- filter(moviesfilteredx, budget > 0)
#Plot graph
plot(rating~budget,data=budgetfiltered ) 
```

Looking at the data as a scatterplot, there seems to be slight positive relationship between the 2 variables.  

To further examine the relationship, I will run a linear regression between the two variablesto determine whether the relationship is significant. The results of the linear regression can be seen below: 

```{r}
#Many rows of our data do not contain a budget, so I will remove those rows 


linregbudget <- lm(rating~budget, data=budgetfiltered)
summary(linregbudget)

plot(rating~budget,data=budgetfiltered ) 
abline(linregbudget, col="red")

#Judging by the high p-value, we can conclude that a films budget does not affect its reception among the audience. 
```
Judging by the linear regression summary, we can conclude that there is a meaningful relationship between the dependant variable, rating, and the independant variable, budget. This is due to the fact that the p-value (2e-16) is significantly smaller than the significance level of 0.05 we are using in this report. 

### How length affects a films reception 

Next, I will be looking at how a film's length affects the reception of a film. 

In order to determine the relationship between the dependant variable (reception) and the independant variable (length), I will run a linear regression, but first I will visualize the data with a scatterplot. 

```{r}
plot(rating~length,data=moviesfilteredx )
```
After observing the scatterplot showing the relationship between reception and length, it is clear that we have an outlier (a film titled 'A Cure for Insomnia' that is over 5000 minutes long — fitting title). I will remove the outlier and perform  linear regression on the data without the outlier. Below, I will display what the data looks like without the outlier. 

```{r}
movieswithoutcureforinsomnia <- filter(moviesfilteredx, length < 3000)
plot(rating~length,data=movieswithoutcureforinsomnia )

```

Although there is a large cluster on the left side of the graph, there does seem to be a slight positive relationship that indicates  a longer length results in a better film rating. To confirm this, we can look at the summary statistics for the linear regression. 

```{r}
#Linear regression between rating and length
linreglength <- lm(rating~length, data=movieswithoutcureforinsomnia)
#Summary statistics for linear regression
summary(linreglength)

plot(rating~length,data=movieswithoutcureforinsomnia )
abline(linreglength, col="red")

```
After observing the summary statistics related to the relationship between rating and length, it is clear that we have a really strong relationship. The p-value of 2e-16 is significantly smaller than our significance level of 0.05 indicating that there is a strong relationship between the variables rating and length. 

### How the year of release affects a films reception 

Last, we will look at how the year a film's released affects it's reception.  

To begin, we will visualize the data with a scatterplot to identify whether we can see a relationship. 

```{r}
plot(rating~year,data=moviesfilteredx )
```

Judging by the scatterplot, there doesn't seem to be an obvious relationship between the two variables. 

We can accurately identify whether there is a relationship by examining the linear regression between the two variables. 

```{r}
linregyear <- lm(rating~year, data=moviesfilteredx)
summary(linregyear)
```

Surprisingly, there is a strong negative relationship. By observing the graph, I thought that there wasn't a significant relationship as I didn't see any blatant positive or negative relationship. But the slightly negative slope with a p-value much smaller than 0.05 indicates that there is a negative relationship between year and reception.  

------------------------------------------------------------------------------------------------------------


## Discussion 

From this results of the analysis, I can conclude that there is a positive relationship between reception and budget, a positive relationshop between reception and length, and a negative relationship between reception and year of release. 

The positive relationship between a film's reception and budget indicates that as a film's budget increases,  the reception of the film increases as well. Many of the film's released before 1950 had positive reception, despite the low budget; this can be a result of the fact that many IMDB reviewers rated these films based on a set of criteria that they may have not used for films released after 1950. This criterion may place a heavy emphasis on what filmmakers were able to do with technology and norms relative to that time. Because of this dichotomy between the seperate eras, I decided to focus exclusively on the era after the 1950s and discovered that there is a positive relationship between a film's reception and budget. 

A larger budget means that the film production has more money to spend; this can results in better actors, special effects, marketing, etc. which can affect how an audience reacts to the film. Despite the fact that a film production may have more money to spend, it doesn't always result in an overall better film. Critics may interpret a good film differently than what an audience interprets as a good film. 

The positive relationshop between a film's reception and length indicates that as a film's length increases, then the reception of the film increases as well. Films released before 1950 were often characterized by their short lengths (especially during film eras before the 1920s). Due to their short lengths and the fact that cinema wasn't used in it's typical way at that time, I decided to focus excluvely on films released after the 1950s. 

Film studios often have a final say before releasing a film. If a film is long, that may advise cutting the film so that it appeals to the masses. For a film to have a long runtime, it would need to be justified; the film would have to be good enough to be worth the audience's time. This may be a reason why a longer run time leads to a better reception. Good filmmakers often want their film to be a result of their uncompromised — without cuts— vision leading to more films that are longer with a better reception.  

The negative relationship between a film's reception and year indicates that older films are likely to have better reception than newer films.  This relationship could be a result of a multitude of factors. As we approach the 2000s, creating a film did not have many barriers as it did many years ago. During the 1950s, to create a film was a lot more expensive due to expensive, heavy, equipment, whereas in the 2000s, film could be be made on a lightweight, cheap video camera.Because of the plight many filmmarkers faced during the 1950s, it may have resulted in a more well-thought out —and therefore, better— films. 

The data was collected from IMDB meaning that a lot of earlier films were rated retroactively. This may result in an inflated score for some of the older films. Many people adopt an 'older is better' mentality when watching older films, which causes them to rate older films with a higher score compared to newer films. 

The results from the analysis can assist people in many ways. When deciding on a good film to watch, the positive relation reception has with budget and length can help identify whether a film may be a good or a bad film; the negative relation between reception and year of release can also aid in helping determine the film's reception. Additionally, when creating a film, some of these relationships can aid in forecasting it's reception. A longer film, with a large budget may result in a film that is well received. 

There are some weaknesses from our analysis, partly due to the dataset used. To measure an audience's reception, we are using a website that wasn't popular until the 21st century. This means that a lot of our reviews for older films are retrospective, which means that may not be the most accurate. Also, because the 



Advantage

Disadvantage
IMDB is a newer site so we are not seeing reviews from people now. 

The results of our analysis do contain weaknesses, many of them stem from the data being used.

## Appendix/References

Show all lines of code and extranneous graphs


